---
title: "Price Prediction of Houses"
author: "Kowshik Kumar B"
date: "11/5/2019"
output: 
  html_document:
      code_folding: show
---

# Calling the Libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
#library(devtools)
#install_github("vqv/ggbiplot")
library(leaps)
library(MASS)
library(e1071) 
library(readr)
library(e1071)
library(Metrics)
library(vioplot)
library(randomForest)
library(ROCR)
```

# Importing and understanding the "houses" data

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
df <- read_csv("houses.csv")
head(df)                        #Accessing few first rows of the "Houses" dataset.
tail(df)                        #Accessing few last rows of the "Houeses" dataset.              
dim(df)                         #Checking the Dimensions of the dataset.
str(df)                        #Checking the structure of the variables of the "Houses" dataset.
```


```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
summary(df)                   #Accessing the Summary of the "Houses" dataset. 
```


# Checking the frequency of the variables of the "Houses" dataset.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
table(df$bedrooms)
table(df$bathrooms)
table(df$waterfront)
table(df$view)
table(df$condition)
table(df$grade)
```

we can see there are majority of 2,3,4,5,6 Bedrooms; 1-5 Bathrooms;No waterfronts; 3-5 condition and 5-10 Grade in the "Houses" dataset.


```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
df$id=NULL
```

# Correlation of the variables

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
library(corrplot)
corrplot(cor(df),method = 'circle',order="hclust")
cor(df)
```

# Splitting the Dataset into Train and Test (60:40) datasets 

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
set.seed(12345)
train.rows = sample(rownames(df), dim(df)[1]*0.6)
train.df = df[train.rows, ]

valid.rows = setdiff(rownames(df), train.rows) 
valid.df = df[valid.rows, ]
```


# Data Visualization of the "Houses" dataset

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
###Code for Additional Clean Graphing of visualization plots###
library("ggplot2")
cleanup = theme(panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.background = element_blank(),
                axis.line.x = element_line(color = 'black'),
                axis.line.y = element_line(color = 'black'),
                legend.key = element_rect(fill = 'white'),
                text = element_text(size = 15))

```

# Scatterplots

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
library(grid)
library(gridExtra)
library("ggplot2")

s1=ggplot(data = train.df, mapping = aes(x =train.df$bedrooms,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Bedrooms ") +
    xlab("Number of Bedrooms")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s2=ggplot(data = train.df, mapping = aes(x =train.df$bathrooms,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Bathrooms ") +
    xlab("Number of Bathrooms")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s3=ggplot(data = train.df, mapping = aes(x =train.df$sqft_living,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Sqft_living ") +
    xlab("Sqft_living")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s4=ggplot(data = train.df, mapping = aes(x =train.df$waterfront,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and waterfront ") +
    xlab("Waterfront")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s5=ggplot(data = train.df, mapping = aes(x =train.df$floors,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Floors ") +
    xlab("Floors")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'grey')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s6=ggplot(data = train.df, mapping = aes(x =train.df$view,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and view ") +
    xlab("view")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s7=ggplot(data = train.df, mapping = aes(x =train.df$view,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and  ") +
    xlab("Number of Bedrooms")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s8=ggplot(data = train.df, mapping = aes(x =train.df$grade,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Grade ") +
    xlab("Grade")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s9=ggplot(data = train.df, mapping = aes(x =train.df$sqft_above,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Sqft_above ") +
    xlab("Sqft_above")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'grey')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s10=ggplot(data = train.df, mapping = aes(x =train.df$sqft_basement,y=train.df$price)) +
    geom_point()+
  ggtitle("Scatterplot of Price and Sqft_basement ") +
    xlab("Sqft_basement")+
    ylab("Price")+
    geom_smooth(method = 'lm', color = 'green1')+
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

s1
s2
s3
s4
s5
s6
s7
s8
s9
s10
```

# Histograms

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
library("ggplot2")
h1=ggplot(data = df, mapping = aes(x = df$price)) +
  geom_histogram( bins=30, fill = "steelblue") +
  labs(x = "Price of the Houses") +
  ggtitle("Histogram of Price of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup
h2=ggplot(data = df, mapping = aes(x = df$bedrooms)) +
  geom_histogram( bins=30, fill = "darkolivegreen") +
  labs(x = "Number of Bedrooms") +
  ggtitle("Histogram of number of Bedrooms of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup
h3=ggplot(data = df, mapping = aes(x = df$bathrooms)) +
  geom_histogram( bins=30, fill = "green1") +
  labs(x = "Number of Bathrooms") +
  ggtitle("Histogram of number of Bathrooms of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup
h4=ggplot(data = df, mapping = aes(x = df$sqft_lot)) +
  geom_histogram( bins=30, fill = "tomato1") +
  labs(x = "Lot Size(Sqft") +
  ggtitle("Histogram of Lot size of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
  cleanup
h5=ggplot(data = df, mapping = aes(x = df$waterfront)) +
  geom_histogram( bins=30, fill = "red1") +
  labs(x = "Waterfront") +
  ggtitle("Histogram of waterfront for Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup
h6=ggplot(data = df, mapping = aes(x = df$yr_built)) +
  geom_histogram( bins=30, fill = "brown1") +
  labs(x = "Built Year") +
  ggtitle("Histogram of Year Built of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup
h7=ggplot(data = df, mapping = aes(x = df$yr_renovated)) +
  geom_histogram( bins=30, fill = "grey1") +
  labs(x = "Renovated Year") +
  ggtitle("Histogram of Year Renovated of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup
h8=ggplot(data = df, mapping = aes(x = df$age)) +
  geom_histogram( bins=30, fill = "orange1") +
  labs(x = "Age of the house") +
  ggtitle("Histogram of age of Houses in King County ") +
  theme(plot.title = element_text(hjust = 0.5)) +
    cleanup

h1
h2
h3
h4
h5
h6
h7
h8
```

Boxplot and kernel density plots of discrete variables of the "Houses" dataset.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
vioplot(df$bedrooms,col="gold")
vioplot(df$bathrooms,col="brown1")
vioplot(df$view,col="yellow1")
vioplot(df$waterfront,col="green1")
vioplot(df$grade,col="tomato1")
vioplot(df$condition,col="steelblue")
```



## Data Modelling

# Linear Model

Using Backward Elimination Wrapper method for the Multi-Linear Regression model

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
full.model=glm(price~.,data=train.df)
summary(full.model)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
step.model <- stepAIC(full.model, direction = "both", trace = TRUE)
summary(step.model)
```


```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
plot(full.model)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
final.model=glm(price ~ bedrooms + bathrooms + sqft_living + sqft_lot + 
    floors + waterfront + view + condition + grade + yr_built + 
    age,data=train.df)
summary(final.model)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
anova(final.model)      #Analysis of variance of the final model
```


# Assumption checking

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
plot(final.model)
```


```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
prediction_glm=predict(final.model,valid.df,type="response")
rmse(valid.df$price,prediction_glm)
```



# Support Vector Machine 

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
model.svm <- svm(price ~ bedrooms + bathrooms + sqft_living + sqft_lot +floors + waterfront + view + condition + grade +yr_built +  age,data = df,kernel="linear")
summary(model.svm)
plot(model.svm)
 prediction_svm = predict(model.svm, valid.df, type="response")
 rmse(valid.df$price,prediction_svm)
```


# Random Forest

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
set.seed(12345)
reg.randomForest=price ~ bedrooms + bathrooms + sqft_living + sqft_lot +floors + waterfront + view + condition + grade +yr_built +  age
 model.randomForest= randomForest(reg.randomForest,df,ntree = 100, nodesize = 10, proximity = T)
 model.randomForest
 plot(model.randomForest)
 varImpPlot(model.randomForest)
 prediction_randomForest = predict(model.randomForest, valid.df, type="response")
 rmse(valid.df$price,prediction_randomForest)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}
library(car)
vif(final.model)
```

